Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-05-31 doc/paper.tex
okay, NOW the common-noun batch is done
hammering out bits of the first section of the introduction

\begin{abstract}
Hello, World!

Here is some normal spacing.

\zh{你好，世界！}

Scalable parallel corpus generation
\end{abstract}

Curiosity: how many AMTA pages are the EMNLP instructions? i.e., how long would my 4-page EMNLP paper be?

\section{Introduction}

RNN encoder-decoders are not the first corpus-based MT methodology (EBMT then SMT then this), and it is unlikely in this era of Big Data that they will be the last.
Focusing on corpus generation allows linguists to focus on capturing the most statistically prevalent linguistic phenomena, while insulating them from the strengths and weaknesses of various implementations.
For instance, EBMT suffers from relatively poor generalization and SMT has problems with nonlocality.
(RNN built on relatively shaky ground, since word vectors don't QUITE make sense for function words?)
Seems somewhat foolhardy to marginalize the domain experts
Significantly more engineering (man-hours?) seems to go into the machine-learning component.  (Well, granted, there are human translators who are spending many man-hours generating sentence pairs. But they're not intended for MT.)
Traditionally, corpus construction has been considered prohibitively expensive, ``scavenged'' from other sources like parliamentary proceedings or lecture subtitles.
But must this be so? Vast untapped resource of students or whatever.

One might conjecture that the fundamental problem in natural language processing is one of data representation

With the ALPAC report's pessimistic outlook on MT arguably responsible for sparking decades of fundamental linguistics research in the tradition of Chomsky, it is only fitting that we now bring linguistics back?
No, my Return of the Linguist had more to do with how RBMT systems have fallen out of favor in research

@@

Return of the linguist: it is relatively easy to come up with examples that industrial-grade systems fail. (hmm,)
he was shown the car
他 给 展示 了 车?

@@

@@

@@

stanford POS tagger (incomplete bibtex entry?) \mycitep{postagger:2003}

@@

In the Moses toolchain, the {\small \tt mkcls} component \mycitep{och:98} performs this process in reverse, binning words into automatically-generated equivalence classes to alleviate data sparsity issues.
% USUALLY alignment only, CAN be applied to translation... http://www.statmt.org/moses/?n=Advanced.Models#ntoc6
With some effort, you can use these bins during steps other than alignment, although this is not done by default.
...
Moreover, as they are upstream of GIZA++, these bins are constructed at the lexical level, not the phrase level, whereas the templates described below can make substitutions at the level of larger constituents and phrases.
Furthermore, it is run at training time and cannot help with unknown words at test time (is this true??)
The only real cure for unknown words is moar data

@@

fire a linguist and accuracy goes up

collobert and weston: learn words from letters. interesting from an AI perspective, seems unlikely to improve performance in the near future

@@

@@

@@

IWSLT: TED talks are a series of lectures, which makes them something intermediate between the conversational language of movie subtitles and the formal written language of UN resolutions.
Should I mention that you can kind of discern this from their POS distributions? I would probably have to mention something concrete then... But yeah, I guess it's worth mentioning if I have space, which I almost certainly do for AMTA

@@

\section{Future Work}

@@

For simplicity, the current implementation is lexically one-to-one; for scalability in the number of languages, you ideally want many-to-many.

@@

Localization of semantic taxonomy - this would have to be one-to-one, and you'd think it would clutter up the raw text UI...

@@

@@

If I really am submitting to this conference too, then I can afford (and probably want to?) draw out the length, 
} % end \amtaonly

@@

>>>>

\begin{abstract}
An increased focus on

is proposed.

Preliminary results do not show much improvement.
This is because we need to scale up?
Future work is proposed.

\end{abstract}

\section{Motivation} % Previously called ``Introduction''

As early as Jelinek's (in)famous 1980's quote, ``{\em Every time I fire a linguist, my performance goes up},'' both linguists and linguistics have been increasingly marginalized in natural language processing. 
More recently, for example, \mynewcite{collobert 2011 footnote 8} mention the possibility of ``{\em [ideally]...to learn from letter sequences rather than words},'' illustrating an AI-motivated desire to move away from even fundamental linguistic concepts.
A specific case of this trend is the gradual shift away from classical linguistic rule-based machine translation (MT) \mycitep{hutchins} to  contemporary machine learning approaches, such as staistical phrase-based systems like Moses \mycitep{moses} and more recent deep learning approaches \mycitep{sutskever, bengio}.

In the end, though, machine learning is learning from data, and even the most bleeding-edge MT system relies crucially on multilingual parallel corpora.  In this paper, we propose recalling linguists from their increasing exile from MT to handle what they know best: language. 
Specifically, we believe that machine translation could benefit from high-quality, ``designer'' parallel corpora that have been hand-crafted by linguists. This could take the form of amplified training sets in the short run, and perhaps a ``universal translation memory'' in the long run.

This division of labor would let linguists be linguists, with MT system development proceeding in parallel, consuming the linguists' work downstream.

UGH I have nothing really concrete to show here... this is like an opinion paper at this point...
Although I guess I do have this stupid system I worked on, and I'll push it to the web, and they can laugh at how pathetic it is.

Significantly more developer-months go into the MT system.\footnote{Of course, hundreds of translator-hours }

niceties like sentence alignment  
(Well, granted, there are human translators who are spending many man-hours generating sentence pairs. But they're not intended for MT.)
There have also been attempts to

@@

I think this first part of the introduction should be short, and then some subsections below?

Traditionally, corpus construction has been considered prohibitively expensive, ``scavenged'' from other sources like parliamentary proceedings or lecture subtitles.
But must this be so? Vast untapped resource of students or whatever.

No, my Return of the Linguist had more to do with how RBMT systems have fallen out of favor in research
banishment has been premature? weren't using the linguist in the right way?

@@

The machine translation task we consider in this paper is the IWSLT shared task

A brief analysis of corpus quality

\mycitep{multeval}

Do I want to display an example bad-sentence pair? 
Some of the lines are duplicated. Examining 
Just mention for now that there is an entire lecture that is bad
Also hand-purged several other sentences

TODO: table of scores after purging

This results in a significant TER score change that is almost as large as the score difference between the highest and lowest ranking system from the evaluation campaign.  
Interestingly, BLEU scores stayed relatively robust.

@@

\subsection{Upsides, Potential Applications}

% interesting, but probably confusing to readers?
%In the language of agile software development, test-driven MT system development, in which linguists pose sentence pairs, and MT developers work to translate them

Focusing on corpus generation would allow linguists to focus their attention on significant linguistic phenomena

methods every few years; linguistic theories, representations every decade or so; parallel corpora live forever

lexical ambiguity is naturally alleviated

RBMT - countless developer-years of expertise

unlike all the different competing methods, it seems that a ``universal translation memory'' would be inherently collaborative (methods are competitive by nature, whereas data can be collaborative)

%, and with the two meeting at the golden railroad spike of the parallel corpus. Apparently I've liked this stupid thing for 10+ years
% on capturing the most statistically prevalent linguistic phenomena, while insulating them from the strengths and weaknesses of various MT implementations. (wait, is that a good thing? don't you need some feedback as to what KIND of corpus the MT system needs?)
%With the ALPAC report's pessimistic outlook on MT arguably responsible for sparking decades of fundamental linguistics research in the tradition of Chomsky, it is only fitting that we now bring linguistics back?
%RNN encoder-decoders are not the first corpus-based MT methodology (EBMT then SMT then this), and it is unlikely in this era of Big Data that they will be the last.
% For instance, EBMT suffers from relatively poor generalization and SMT has problems with nonlocality.
% (RNN built on relatively shaky ground, since word vectors don't QUITE make sense for function words?)
% Seems somewhat foolhardy to marginalize the domain experts (no, lots of deep learning advances have been made by IGNORING expert-created representations)
% One might conjecture that the fundamental problem in natural language processing is one of data representation
%Return of the linguist: it is relatively easy to come up with examples that industrial-grade systems fail. (hmm,)
%he was shown the car \zh{他 给 展示 了 车}?

@@

@@

In the Moses toolchain, the {\small \tt mkcls} component \mycitep{och:98} performs this process in reverse, binning words into automatically-generated equivalence classes to alleviate data sparsity issues.
% USUALLY alignment only, CAN be applied to translation... http://www.statmt.org/moses/?n=Advanced.Models#ntoc6
% uh, so do people already do this??
With some effort, you can use these bins during steps other than alignment, although this is not done by default.
...
Moreover, as they are upstream of GIZA++, these bins are constructed at the lexical level, not the phrase level, whereas the templates described below can make substitutions at the level of larger constituents and phrases.
Furthermore, it is run at training time and cannot help with unknown words at test time (is this true??)
The only real cure for unknown words is moar data

@@

@@

used for corpus analysis stanford POS tagger (incomplete bibtex entry?) \mycitep{postagger:2003}

@@

Out of domain corpus

@@

The corpus, which is freely available, consists of multilingual subtitles of TED talks, which are prepared lectures on a wide variety of subjects.

IWSLT: TED talks are a series of lectures, which makes them something intermediate between the conversational language of movie subtitles and the formal written language of UN resolutions.

@@

\section{Proposed Future Work}

@@

@@

crowdsourcing for ``amateur linguists''

estimate about 1 noun per minute (my pathetic work speed)

@@

In the short term
Increasing lexical variation by expanding the lexicon, especially for uncommon words in the training corpus, to alleviate data sparsity.
Increasing syntactic variation in a high-quality way by harvesting more templates, in the old EBMT tradition
Work along these lines is currently underway.

@@

If I really am submitting to this conference too, then I can afford (and probably want to?) draw out the length,

Should I mention that you can kind of discern differences between IWSLT, movie subtitles, and MultiUN from their POS distributions? I would probably have to mention something concrete then...? 
} % end \amtaonly

@@

wait, the HCOMP ``short paper'' needs to be 2 pages at MOST, including references
